---
title: Tips in R studio
author: Sikai Huang
date: '2022-05-06'
categories:
  - R
tags:
  - RStudio
slug: r-shortcut
---

## Shortcuts

**R shortcuts** make the programming life easier. Full information is available directly in the ***Navigation Bar - Tools- Keyboard shortcuts help***. I listed commonly used shortcuts that might help **R beginners (students)** like me:

|                              |   Windows & Linux    |          Mac          |
|:-----------------------------|:--------------------:|:---------------------:|
| Open Document                |       Ctrl + O       |        Cmd + O        |
| New R Script                 |   Ctrl + Shift + N   |    Cmd + Shift + N    |
| Run current Line             |       Ctrl + ↵       |        Cmd + ↵        |
| Undo                         |       Ctrl + Z       |        Cmd + Z        |
| Redo                         |   Ctrl + Shift + Z   |    Cmd + Shift + Z    |
| Copy                         |       Ctrl + C       |        Cmd + C        |
| Paste                        |       Ctrl + Y       |        Cmd + Y        |
| Cut                          |       Ctrl + X       |        Cmd + X        |
| Move Lines Up/Down           |     Alt+Up/Down      |    Option+Up/Down     |
| Delete Line                  |       Ctrl + D       |        Cmd + D        |
| Move to Line start           |     Ctrl + Left      |      Cmd + Left       |
| Move to Line end             |     Ctrl + Right     |      Cmd + Right      |
| Select to Line Start         | Ctrl + Shift + Left  | Cmd + + Shift + Left  |
| Select to Line End           | Ctrl + Shift + Right | Cmd + + Shift + Right |
| Insert Pipe Operator( %\>% ) |   Ctrl + Shift + M   |    Cmd + Shift + M    |
| Find                         |       Ctrl + F       |        Cmd + F        |

## Change a value in a specific column according to the value in a specific column
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(DT)
datatable(mtcars, options = list(pageLength = 3))
```

Try to assign ***NA*** to **disp** if **cyl** is equal to **4**. "Mutate" or "If" is overkill in this case.

```{r message=FALSE, warning=FALSE}
mtcars$disp[mtcars$cyl == 4] <- "NA"
datatable(mtcars, options = list(pageLength = 3))
```




