---
title: Tidyverse Dplyr Package
author: Sikai Huang
date: '2022-05-08'
categories:
  - R
slug: dplyr-package
output:
  blogdown::html_page:
    toc: true
    toc_depth: 2
---

# Introduction to dplyr

## Essential package

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(gapminder) # a built-in dataset called "gapminder" in R studio
```

## Filter

Extracting a subset from observations, only containing the row of chosen variables.

```{r, collapse=TRUE}
gapminder %>%
  filter(year == 2002, country == "China") 
```

## Arrange

Sortng observations in ascending or descending order of a particular variable.

```{r}
# Sort in descending order of lifeExp
gapminder %>%
  arrange(desc(lifeExp))  # the example of ascending order is the default setup not shown here
```

## Mutate

Using mutate to change(overwrite an existing one) or create a column

```{r}
# Use mutate to change the lifeExp
gapminder %>%
  mutate(lifeExp = lifeExp * 12)
# Use mutate to create a new column called lifeExpMonths
gapminder %>%
  mutate(lifeExpMonths = lifeExp * 12)
```

## Filter_Arrange_Mutate

```{r}
# Filter, mutate, and arrange the gapminder dataset
gapminder %>%
  filter(country == "Japan") %>%
  mutate(lifeExpMonths = lifeExp * 12) %>%
  arrange(desc(lifeExpMonths))
```

# Grouping and Summarising

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(gapminder)
```

## Summarising

```{r}
gapminder %>%
  filter(year == 2007) %>%
  summarize(medianLifeExp = median(lifeExp, na.rm = TRUE),
            maxGdpPercap = max(gdpPercap, na.rm = TRUE))
```

## Grouping

```{r}
gapminder %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarize(meanLifeExp = mean(lifeExp),
            totalPop = sum(pop)) # yield a value to each continent in a total of 5
```

## Visualizing medianGdpPercap within each continent within each year

```{r message=FALSE, warning=FALSE}
by_year_continent <- gapminder %>%
  group_by(continent, year) %>%
  summarize(medianGdpPercap = median(gdpPercap))
```

```{r}
# Plot the change in medianGdpPercap in each continent over time
library(ggplot2)
ggplot(by_year_continent, aes(x = year, y = medianGdpPercap, color = continent)) +
  geom_point() +
  expand_limits(y = 0) # To make sure the plot's y-axis includes zero
```

# Data visualisation

## Essential package

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(gapminder)
library(ggplot2)
```

## Scatter plot

```{r}
# Create gapminder_1952
gapminder_1952 <- gapminder %>%
  filter(year == 1952)
# Create a scatter plot with pop on the x-axis and lifeExp on the y-axis
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
  geom_point()
```

Some countries have **much more** population than others where we need to scale the data using **scale_x\_log10()**.

```{r}
gapminder_1952 <- gapminder %>%
  filter(year == 1952)
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
  geom_point() +
  scale_x_log10()
```

## Faceting

To give a graph to each categorical variable, divide data into subplots

```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent, size = pop))+
  geom_point() +
  scale_x_log10() +
  facet_wrap(~ year) # "~ year" means by year
```

## Line plot

```{r message=FALSE, warning=FALSE}
# Create a line plot showing the change in medianGdpPercap by continent over time
by_year_continent <- gapminder %>%
  group_by(year, continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

ggplot(by_year_continent, aes(x = year, y = medianGdpPercap, color = continent)) +
  geom_line() +
  expand_limits(y = 0)

```

## Bar plot

```{r}
# Create a bar plot showing medianGdp by continent
by_continent <- gapminder %>%
  filter(year == 1952) %>%
  group_by(continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

ggplot(by_continent, aes(x = continent, y = medianGdpPercap, color = continent)) +
  geom_col()
```

## Histogram

```{r message=FALSE, warning=FALSE}
# Create a histogram of population (pop), with x on a log scale
gapminder_1952 <- gapminder %>%
  filter(year == 1952)
ggplot(gapminder_1952, aes(x = pop))+
  geom_histogram()+
  scale_x_log10()
```

## Box plot

```{r}
# Create a boxplot to compare GDP per capita across continents
gapminder_1957 <- gapminder %>%
  filter(year == 1957)

ggplot(gapminder_1957, aes(x = continent, y = gdpPercap)) +
  geom_boxplot() +
  scale_y_log10()
```

## Asthetics

### Size and Color

```{r}
ggplot(gapminder_1952, aes(x = gdpPercap, y = lifeExp, color = continent, size = pop)) + 
  geom_point() + 
  scale_x_log10()
```

### Title

```{r}
# Add a title to this graph: "Comparing GDP per capita across continents"
ggplot(gapminder_1952, aes(x = continent, y = gdpPercap)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(title = "Comparing GDP per capita across continents", x = "where continent", y = "how much GDP per capita") # Or plot + ggtitle("Comparing GDP per capita across continents") + xlab("where continent") + ylab("how much gdpPercap")
```

### Legend label by x-axis variable

```{r}
ggplot(gapminder_1952, aes(x = continent, y = gdpPercap, fill = continent)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(title = "Comparing GDP per capita across continents", x = "where continent", y = "how much GDP per capita", fill = "Continent")
```
